Hurricane Evacuation Strategies for the State of Mississippi

### 2015-11-16 manuscript.tex
waiting for pull

\section{Markov Process Analysis of Strictly Northern Evacuations}
\label{sec:markov}
  Because our analysis of historical hurricane trends concluded that the vast majority of landfalls occur in the three southern-most counties, the first evacuation strategy we tested was one where all evacuees move strictly northward.
  \subsection{Model Description}
    Given that we have a network structure and associated population flow rates, this problem lends itself well to a Markov process analysis. We can model our evacuation strategy by converting the network of Mississippi counties into a directed graph. For our northward evacuation, this means only drawing highway-based edges to destination nodes with higher latitudes:
    \begin{figure}[H]
      \centering
      \includegraphics[width=.5\textwidth]{figures/full_directed_NS.pdf}
      \caption{Map of Northbound Highway Routes in the State of Mississippi \cite{ms_highways}}
    \end{figure}
    The thickness of each arrow in this graph represents the transition probability from one node to another. In evacuation terms, these are the proportions of a county's population that travel to a destination node during a single update. In order We standardize the time interval of an update to be one hour, and track the number of people in every county after each update in the 4 day period before landfall.

  \subsection{Parameter Values and Justification}
    Using our earlier values for the maximum flow rate of an edge, we can now use this 
  \subsection{Results}
  \subsection{Strengths and Weaknesses}

@@

\section{Stochastic Model of Landfall-Avodiant Evacuations}
\label{sec:stochastic}
  \subsection{Model Description}
    \par In this model, instead of giving citizens particular evacuation plan, we give them a simple rule: move directly away from the hurricane landfall if possible, else move North or North-West. Using the statewide broadcast system, the state of Mississippi can tell its citizens where the expected landfall will be, and which counties will be the most affected. In which case, individual citizens are instructed to leave the counties that will be flooded, and to do so as outlined above. This means less coordination from the state, and more options for each citizens. At the cost of not having designated areas, and therefore designated refuge areas for citizens, evacuation is faster because it involves less per-county organization.\\
    \par As opposed to the model described in \nameref{sec:maxflow}, we also allow citizens to leave the state of Mississippi. For example, Arkansas is less likely to be hit by the full force of hurricanes that move through Mississippi \cite{5news}, as such, it is a good place to take refuge while the storm passes. In order to achieve this, we add ``unofficial'' nodes in the graph representation of Mississippi. Figure~\ref{fig:highway_map} shows that Hancock and Jackson Counties have roads that lead into Louisiana and Alabama respectively, both of which are potential routes away from the Hurricane's landfall, meaning that such nodes are important when movement can also be lateral. Because we are only interested in the movement of citizens within Mississippi, we can assume that citizens that reach the out-of-state nodes keep moving further and further away from the hurricane, though we are not tracking them. We created four out-of-state nodes North, South, East and West of Mississippi, each connected in a way that satisfies the highways mapped in Figure~\ref{fig:highway_map}.
  \subsection{Parameter Values and Justification}
    \par The parameters to this model are the same as the parameters for the models outlined above, except that this model uses the complete highway map of Mississippi, instead of only considering vertical highways. This algorithm expects three inputs outside of the parameters: the first, less accurate estimate of the affected counties, the more accurate second estimate and the final, most accurate estimate. Notice that as described, these inputs actually abstract away the hurricane's category. Instead, the category is reflected by the size of the list of the affected counties given as the respective inputs for the estimates.
  \subsection{Results}
    \begin{figure}
      \center
      \begin{subfigure}[b]{0.5\textwidth}
        \center
        \includegraphics[width=\linewidth]{figures/pred_hancock-crop.pdf}
        \caption{Predicted Evacuation for Landfall in Hancock County}
      \end{subfigure}~
      \begin{subfigure}[b]{0.5\textwidth}
        \center
        \includegraphics[width=\linewidth]{figures/pred_jackson-crop.pdf}
        \caption{Predicted Evacuation for Landfall in Jackson County}
      \end{subfigure}
      \caption{}
    \end{figure}
  \subsection{Strengths and Weaknesses}

>>>>

\section{Markov Process Analysis of Strictly Northern Evacuations}
\label{sec:markov}
  Because our analysis of historical hurricane trends concluded that the vast majority of landfalls occur in the three southern-most counties, the first evacuation strategy we tested was one where all evacuees move strictly northward.
  \subsection{Model Description}
    Given that we have a network structure and associated population flow rates, this problem lends itself well to a Markov process analysis. We can model our evacuation strategy by converting the network of Mississippi counties into a directed graph. For our northward evacuation, this means only drawing highway edges to destination nodes with higher latitudes:
    \begin{figure}[H]
      \centering
      \includegraphics[width=.5\textwidth]{full_directed_ns.pdf}
      \caption{Map of Major Highways in the State of Mississippi \cite{ms_highways}}
    \end{figure}
    Then each edge transition probability in this graph determines the proportion of a county's population that travels to a destination node during a single update. We standardize the time interval of an update to be one hour, and track the number of people in every county after each update in the 4 day period before landfall.

  \subsection{Parameter Values and Justification}
  \subsection{Results}
  \subsection{Strengths and Weaknesses}

@@

\section{Stochastic Model of Landfall-Avodiant Evacuations}
\label{sec:stochastic}
  \subsection{Model Description}
    \par In this model, instead of giving citizens particular evacuation plan, we give them a simple rule: move directly away from the hurricane landfall if possible, else move North or North-West. Using the statewide broadcast system, the state of Mississippi can tell its citizens where the expected landfall will be, and which counties will be the most affected. In which case, individual citizens are instructed to leave the counties that will be flooded, and to do so as outlined above. This means less coordination from the state, and more options for each citizens. At the cost of not having designated areas, and therefore designated refuge areas for citizens, evacuation is faster because it involves less per-county organization.\\
    \par As opposed to the model described in \nameref{sec:maxflow}, we also allow citizens to leave the state of Mississippi. For example, Arkansas is less likely to be hit by the full force of hurricanes that move through Mississippi \cite{5news}, as such, it is a good place to take refuge while the storm passes. In order to achieve this, we add ``unofficial'' nodes in the graph representation of Mississippi. Figure~\ref{fig:highway_map} shows that Hancock and Jackson Counties have roads that lead into Louisiana and Alabama respectively, both of which are potential routes away from the Hurricane's landfall, meaning that such nodes are important when movement can also be lateral. Because we are only interested in the movement of citizens within Mississippi, we can assume that citizens that reach the out-of-state nodes keep moving further and further away from the hurricane, though we are not tracking them. We created four out-of-state nodes North, South, East and West of Mississippi, each connected in a way that satisfies the highways mapped in Figure~\ref{fig:highway_map}.
  \subsection{Parameter Values and Justification}
    \par The parameters to this model are the same as the parameters for the models outlined above, except that this model uses the complete highway map of Mississippi, instead of only considering vertical highways. This algorithm expects three inputs outside of the parameters: the first, less accurate estimate of the affected counties, the more accurate second estimate and the final, most accurate estimate. Notice that as described, these inputs actually abstract away the hurricane's category. Instead, the category is reflected by the size of the list of the affected counties given as the respective inputs for the estimates. In order to generate these input, we used the historical hurricane data. Given the landfall and the hurricane's category, we can generate three sets of decreasing size containing neighboring counties. We can then run the simulation on these inputs.
  \subsection{Results}
    This model's main goal is to find the amount of time it takes to evacuate the affected states.
    \begin{figure}
      \center
      \begin{subfigure}[b]{0.5\textwidth}
        \center
        \includegraphics[width=\linewidth]{figures/pred_hancock-crop.pdf}
        \caption{Predicted Evacuation for Landfall in Hancock County}
      \end{subfigure}~
      \begin{subfigure}[b]{0.5\textwidth}
        \center
        \includegraphics[width=\linewidth]{figures/pred_jackson-crop.pdf}
        \caption{Predicted Evacuation for Landfall in Jackson County}
      \end{subfigure}
      \caption{}
      \label{}
    \end{figure}
  \subsection{Strengths and Weaknesses}
